---
title: "smartphone_quality_markdown"
output:  github_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Aim of this project is to use ordinal logistic regression to look for associations between data quality and smartphone characteristics (age, storage, ram etc).

Libraries: 
```{r load_library, results='hide', message=F, warning=F}
library(readr)
library(dplyr)
library(ordinal)
library(tidymodels)
```


Load the data:
```{r load_data, results='hide', message=F, warning=F}
combined_data <- read_csv("../combined_data_version2.csv")
```

The variables: 

* cores - how many cores in the CPU (used for parallel computing, it is good for opening several apps at the same time)

* storageX - disk size, X different storages depending on model etc. 

* ramX - memory ram  (it measures the smartphone capacity for handling multiple tasks)

* battery capacity - this is the physical battery capacity at manufacture measured in mAh, this decreases with the usage so we don't really know the actual capacity when the patients got the app

* GHz_min and GHz_max - processor frequency (how strong is the cpu)

* release_year: should correlate with age of phone (transform to years since release)

* quality : main outcome, GOOD, BAD, MEDIUM

* foreground_app_version : bd4qol app version  (categorical)

* os_version : version of android os (categorical)

* sdk: sdk app version (categorical)

* brand: phone brand

## Remove duplicates
Duplicates: multiple rows/patient ids. Will have price, can remove some based on currency (only keep euro/pounds). Else, try version keeping duplicates and compare to case where all duplicates are removed. Make 2 dataframes later

```{r duplicates_country}
# Dataframe with only unique patient_ids
unique_patients <- combined_data %>%
  group_by(patient_id) %>%
  filter(n() == 1) %>%
  ungroup()

# Dataframe with duplicate patient_ids, filtered to include 'EUR' in price
duplicate_patients_eur <- combined_data %>%
  group_by(patient_id) %>%
  filter(n() > 1 & grepl("EUR", price)) %>%
  ungroup()

# Combining both dataframes
combined_data <- bind_rows(unique_patients, duplicate_patients_eur)

```


## Explore and transform data:

Quality
```{r expl_quality}
# Fix the order
combined_data <- combined_data %>% mutate(quality = factor(quality,
                                                           levels = c("BAD", "MEDIUM", "GOOD"))
                                          )
table(combined_data$quality)
```

cores: most are 8 cores
```{r expl_cores}
table(combined_data$cores)
hist(combined_data$cores)
```

battery cap:
```{r expl_bat}
hist(combined_data$battery_capacity)
```

release_year: transform to years since release (2024-release_year)
```{r expl_year}
table(combined_data$release_year)
combined_data <- combined_data %>% mutate(years_since_release = 2024-release_year)
hist(combined_data$years_since_release)

```

foreground_app_version, brand, os_version, sdk: make into factors
```{r expl_categories}
combined_data <- combined_data %>% mutate(foreground_app_version = factor(foreground_app_version),
                                          brand = factor(brand),
                                          os_version = factor(os_version),
                                          sdk = factor(sdk))

```

GHz_min and GHz_max: keep as is

storage: can have different amounts depending on model; make a max and min storage
```{r expl_storage}
combined_data <- combined_data %>%
  rowwise() %>%
  mutate(
    storage_min = min(c(storage1, storage2, storage3, storage4, storage5, storage6, storage7, storage8, storage9),na.rm=T), 
    storage_max = max(c(storage1, storage2, storage3, storage4, storage5, storage6, storage7, storage8, storage9),na.rm=T)
  ) %>%
  ungroup()

hist(combined_data$storage_min)
hist(combined_data$storage_max)

```

RAM: can depend on the model. Make a max, min, median, a binary (<5 across min_max), a binary <5 median
```{r expl_ram}
combined_data <- combined_data %>%
  rowwise() %>%
  mutate(
    ram_min = min(c(ram1, ram2, ram3, ram4, ram5, ram6, ram7, ram8, ram9),na.rm=T), 
    ram_max = max(c(ram1, ram2, ram3, ram4, ram5, ram6, ram7, ram8, ram9),na.rm=T),
    ram_median = median(c(ram1, ram2, ram3, ram4, ram5, ram6, ram7, ram8, ram9),na.rm=T)
  ) %>%
  ungroup()

hist(combined_data$ram_min)
hist(combined_data$ram_max)
hist(combined_data$ram_median)

combined_data <- combined_data %>%
  mutate(
    ram_bin_minmax_larger_then_5 = if_else(ram_max > 5, true = 1,false = 0), 
    ram_bin_median_larger_then_5 = if_else(ram_median > 5, true = 1,false = 0)
  )

table(combined_data$ram_bin_minmax_larger_then_5)
table(combined_data$ram_bin_median_larger_then_5)

```

Make smaller dataset with only relevant variables & make no_duplicates dataset:
```{r reduced_dataset}
data_reduced <- combined_data %>% dplyr::select(c(patient_id, quality, cores, battery_capacity, foreground_app_version,
                                           os_version, sdk, brand, GHz_max, GHz_min, years_since_release, storage_min,
                                           storage_max, ram_min, ram_max, ram_median, ram_bin_minmax_larger_then_5, ram_bin_median_larger_then_5))

data_reduced_noDups <- data_reduced %>%
  group_by(patient_id) %>%
  filter(n() == 1) %>%
  ungroup()

```

## Analyses: with duplicates

Ordered logistic regression (proportional odds logistic regression): too many parameters, remove brand, sdk, os_version, foreground. Also scaled all numeric variables.
```{r ordered_log_reg}
mod_ord_logreg1_max <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced)
summary(mod_ord_logreg1_max)

mod_ord_logreg1_min <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_min) + scale(storage_min) + scale(ram_min), data = data_reduced)
summary(mod_ord_logreg1_min)

mod_ord_logreg1_median <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_min) + scale(storage_min) + scale(ram_median), data = data_reduced)
summary(mod_ord_logreg1_median)

```
Nothing found here. 

Try univariate models:
```{r ordered_log_reg_univar}
summary(clm(quality ~ scale(cores), data = data_reduced)) #almost sign
summary(clm(quality ~ scale(battery_capacity), data = data_reduced))
summary(clm(quality ~ scale(years_since_release) , data = data_reduced))
summary(clm(quality ~ scale(GHz_max), data = data_reduced)) # sign
summary(clm(quality ~ scale(GHz_min), data = data_reduced))
summary(clm(quality ~ scale(storage_max), data = data_reduced)) # sign
summary(clm(quality ~ scale(storage_min), data = data_reduced)) # sign
summary(clm(quality ~ scale(ram_max), data = data_reduced)) #sign
summary(clm(quality ~ scale(ram_min), data = data_reduced)) #sign
summary(clm(quality ~ scale(ram_median), data = data_reduced)) #sign
summary(clm(quality ~ scale(ram_bin_minmax_larger_then_5), data = data_reduced)) #sign
summary(clm(quality ~ scale(ram_bin_median_larger_then_5), data = data_reduced)) #sign
summary(clm(quality ~ sdk, data = data_reduced))
summary(clm(quality ~ os_version, data = data_reduced)) #sign
summary(clm(quality ~ foreground_app_version, data = data_reduced))


```

Try logistic regression where outcome is binary: good (1) or not good (0)
```{r logistic_reg_good}
data_reduced <- data_reduced %>% mutate(quality_good = if_else(quality == "GOOD", true = 1, false = 0))
summary(glm(quality_good ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced, family=binomial()))

```

Try logistic regression where outcome is binary: bad (1) or not bad (0)
```{r logistic_reg_bad}
data_reduced <- data_reduced %>% mutate(quality_bad = if_else(quality == "BAD", true = 1, false = 0))
summary(glm(quality_bad ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced, family=binomial()))

```


## Analyses: without duplicates

Ordered logistic regression (proportional odds logistic regression): too many parameters, remove brand, sdk, os_version, foreground. Also scaled all numeric variables.
```{r ordered_log_reg_noDups}
mod_ord_logreg1_max_noDups <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced_noDups)
summary(mod_ord_logreg1_max_noDups)

mod_ord_logreg1_min_noDups <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_min) + scale(storage_min) + scale(ram_min), data = data_reduced_noDups)
summary(mod_ord_logreg1_min_noDups)

mod_ord_logreg1_median_noDups <- clm(quality ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_min) + scale(storage_min) + scale(ram_median), data = data_reduced_noDups)
summary(mod_ord_logreg1_median_noDups)

```
Nothing found here. 

Try univariate models:
```{r ordered_log_reg_univar_noDups}
summary(clm(quality ~ scale(cores), data = data_reduced_noDups)) #sign
summary(clm(quality ~ scale(battery_capacity), data = data_reduced_noDups))
summary(clm(quality ~ scale(years_since_release) , data = data_reduced_noDups))
summary(clm(quality ~ scale(GHz_max), data = data_reduced_noDups)) # sign
summary(clm(quality ~ scale(GHz_min), data = data_reduced_noDups))
summary(clm(quality ~ scale(storage_max), data = data_reduced_noDups)) # sign
summary(clm(quality ~ scale(storage_min), data = data_reduced_noDups)) # sign
summary(clm(quality ~ scale(ram_max), data = data_reduced_noDups)) #sign
summary(clm(quality ~ scale(ram_min), data = data_reduced_noDups)) #almost sign
summary(clm(quality ~ scale(ram_median), data = data_reduced_noDups)) #sign
summary(clm(quality ~ scale(ram_bin_minmax_larger_then_5), data = data_reduced_noDups)) #sign
summary(clm(quality ~ scale(ram_bin_median_larger_then_5), data = data_reduced_noDups)) #almost sign
summary(clm(quality ~ os_version, data = data_reduced_noDups))

```

Try logistic regression where outcome is binary: good (1) or not good (0)
```{r logistic_reg_good_noDups}
data_reduced_noDups <- data_reduced_noDups %>% mutate(quality_good = if_else(quality == "GOOD", true = 1, false = 0))
summary(glm(quality_good ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced_noDups, family=binomial()))

```

Try logistic regression where outcome is binary: bad (1) or not bad (0)
```{r logistic_reg_bad_noDups}
data_reduced_noDups <- data_reduced_noDups %>% mutate(quality_bad = if_else(quality == "BAD", true = 1, false = 0))
summary(glm(quality_bad ~ scale(cores) + scale(battery_capacity) + scale(years_since_release) +
                          scale(GHz_max) + scale(storage_max) + scale(ram_max), data = data_reduced_noDups, family=binomial()))

```

## Make plots to show effects
From the analyses it seems that ram and storage are the main ones with cores as a maybe, but not able to specify which one specifically

```{r plots}
par(mfrow=c(1,3))
sciplot::lineplot.CI(quality, ram_median, data = data_reduced)
sciplot::lineplot.CI(quality, storage_min, data = data_reduced)
sciplot::lineplot.CI(quality, cores, data = data_reduced)
par(mfrow=c(1,1))


```





